---
title: "Progress Report"
author: "Victor Huang"
date: "November 17, 2020"
output: 
  pdf_document:
    latex_engine: pdflatex
---


    

```{r setup, echo=FALSE,warning=FALSE,message=FALSE}

source(paste0(here::here(),"/functions/set_para.R", ''))
source(paste0(here::here(),"/functions/pop_gen.R", ''))
source(paste0(here::here(),"/functions/pop_draw.R", ''))
source(paste0(here::here(),"/functions/simulation2.R", ''))
library(tidyverse)
library(kableExtra)
```

## Overview:
  Concerned variables and underlying Monte-Carlo model:  
    
    nSims: Number of simulations to run for each groups of parameter.  
    N: Size of population.  
    p: Probablity an observation is sampled.  
    sigma_e2: Variance of Model error e.  
    sigma_epi2: variance of unpredicted error epislon.  
    yi: Dependent variable without unpredicted error.  
    xi: Independent variable.  
    cap_y: Dependent variable with unpredicted error.
      
    Simulation2: A function that randomly draws samples from population given. Each observation \\
    has probability p of being samapled. It repeats this operation nSims times and will output a \\
    vector of beta, the empirical value for an OLS fitting model based on cap_y and xi.  
      
```{r, echo=FALSE}
start.time<-Sys.time()
```
        
          
```{r, results='asis',warning=FALSE,message=FALSE}
# setting parameters
sigma_e2<-1
N<-10000
beta<-2
nSims<-10000

p<-c(0.01,0.1,0.5,0.9,0.99)
sigma_epi2<-c(0,1,2)

x_para<-set_para(0,1)
e_para<-set_para(0,sigma_e2)

# generating population
xi<-pop_gen(x_para,N,'uniform')
ei<-pop_gen(e_para,N,'normal')

yi<-xi*beta+ei



result<-NULL

# for each group of simulations, generate one population with 
# independent error epislon.

for (j in c(1:length(sigma_epi2))){
    epi_para<- set_para(0,sigma_epi2[j])
    epi<-pop_gen(epi_para,N,'normal')
    
    cap_y<-yi+epi
    
# sample from the population with probability p and return the 
# variance of the beta.
  for (i in c(1:length(p))){
    start.time.small<-Sys.time()
    var_beta<-var(simulation2(cap_y,xi,p[i],nSims))
    theo_var<-sigma_epi2[j]/(sum(xi**2))+((1-p[i])/p[i])*
      (sum(xi**2*ei**2))/((sum(xi**2))**2)
    
    end.time.small<-Sys.time()
    
# rearranging the data for presentation and making graph
    out<-cbind(obs=var_beta,theo=theo_var,prob=sprintf('p = %f ',p[i]),
               sigma=sprintf('sigma_epi2 = %f',sigma_epi2[j]),
               time=end.time.small-start.time.small)
    result<-rbind(result,out)
  }
}

```

```{r Plot,include=TRUE}
result1<-as.data.frame(result) %>% gather(type,num,obs:theo) %>% type_convert()

for (i in (c(1:length(p)))){
a<-ggplot(result1 %>% filter(prob==sprintf('p = %f',p[i])))+
    geom_col(aes(sigma,num,fill=type),position='dodge',width = 0.9)+
  geom_text(aes(sigma,num,label=round(as.numeric(num),5),group=type),
            position = position_dodge(width = 0.9))+
  theme(axis.title.x=element_blank(),
        axis.ticks.x=element_blank(),
        plot.margin = unit(c(1,1,1,1), "cm"))+
  ggtitle(sprintf('Observation with p= %f',p[i]))
plot(a)
}

```

```{r Plot2}
for (i in (c(1:length(sigma_epi2)))){
a<-ggplot(result1 %>% filter(sigma==sprintf('sigma_epi2 = %f',sigma_epi2[i])))+
    geom_col(aes(prob,num,fill=type),position='dodge',width = 0.9)+
  geom_text(aes(prob,num,label=round(as.numeric(num),5),group=type),
            position = position_dodge(width = 0.9))+
  theme(axis.title.x=element_blank(),
        axis.ticks.x=element_blank(),
        plot.margin = unit(c(1,1,1,1), "cm"))+
  ggtitle(sprintf('Observation with sigma= %f',sigma_epi2[i]))
plot(a)
}

```

```{r, echo=FALSE}
end.time<-Sys.time()
```

```{r,echo=FALSE,results="asis"}
kable(as.data.frame(result) %>% select(prob,sigma,obs,theo))

kable(rbind(as.data.frame(result) %>% select(prob,sigma,time),
c('time in total','time in total',end.time-start.time)))
```


